Dewlap color variation in \textit{Anolis sagrei} is maintained between habitats within islands of the West Indies

### 2020-08-02 ms/sections/discussion.tex
eod

\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence. Simple models of genetic drift are ruled out and we don’t think it is plasticity. However, our primary test for adaptation—looking for parallel responses across islands—also was falsified. We suggest, however, that selection could be involved, and suggest several ways this might have happened.}

@@

\paragraph{No conclusive evidence for local adaptation of dewlap coloration} One of the most informative tests for adaptation is the convergence of differentiation patterns across replicate islands or localities \citep{Losos2009, Losos2011}. Previous studies have described convergent patterns of dewlap color evolution in similar environments across islands and species \citep{Thorpe2002a, Thorpe2002b}. Others have suggested that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e. UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{A. cristatellus} and \textit{A. cooki} from Puerto Rico \citep{Leal2002, Leal2004}. Although UV reflectance was commonly involved in between-habitat divergence in \textit{A. cooki} and \textit{A. cristatellus}, we found no such patterns in \textit{A. sagrei}, where instead, we found the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}. The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone, and the test of convergence across islands is rejected. Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}. Habitats on different islands may also differ in aspects other than light conditions, such as densities of predators or congeners, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}. In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A. sagrei} on islands with more coexisting species of anoles. Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A. sagrei}.

\paragraph{Sexual selection could be at play} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Andersson1994}. This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g. plasticity or random drift). Substantial levels of promiscuity in \textit{A. sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario. However, \citet{Baeckens2018} found no link between \textit{A. sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos. Another form of sexual selection is the "good genes" model, where the cue under sexual selection is an indicator of individual quality (e.g. better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}. Several studies suggest that this possibility could be the case for anoles. For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A. brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition. \citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A. sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health. Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}. If dewlap coloration is indeed an indicator of male quality in \textit{A. sagrei}, however, this would explain the observed patterns only if cues associated with good genes differ between sites, which is unlikely as it implies an association of male fitness with different dewlap coloration in different habitats, the evolution of which may be counteracted by gene flow (citation to say that if gene flow counters the divergence of one trait, then it is even worse for the association of two traits). Basically, one would expect good genes-sexual selection to favor the same coloration across the archipelago, and would therefore not result in local differences between habitats.

@@

\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors such as parasite load, as mentioned above \citep{Cook2013}. The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}. Animals lack the ability to synthesize carotenoids, and those must therefore be found in the diet, while pterins are synthesized from nucleotides \citep{Goodwin1984, Hill2002, Hill2006}. However, experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A. sagrei} \citep{Steffen2010} nor in \textit{A. distichus} \citep{Ng2013}, which also has an orange-based dewlap. This makes a plastic response to differences in diet across habitats unlikely. Furthermore, developmental plasticity during the ontogeny is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}. The differences we observed could therefore be genetically based. This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A. sagrei}. Response to parasite load is definitely an example of plasticity!\\ % playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore

@@

\paragraph{Implications in the context of speciation} Local adaptation can be a precursor to ecological speciation, a process that may have given rise to the adaptive radiation of \textit{Anolis} lizards \citep{Harmon2003, Gavrilets2009}. Ecologically-mediated divergence of a sexual signal may be a potent path to the evolution of reproductive isolation through divergent sexual selection \citep{Reynolds2007, Servedio2011}. Evidence suggests that dewlap coloration could take this role in anoles \citep{Ng2011, Lambert2013, Geneva2015, Ng2017}, or at least that it is frequently involved in species recognition \citep{Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013, Ingram2016, Baeckens2018}. Although a correlation between dewlap coloration and reproductive isolation is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular is capable of rapid macroevolution throughout its macroevolutionary history \citep{Nicholson2007}. For example, \textit{A. conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}. We present evidence of multiple cases of within-island dewlap divergence over small geographical scales in \textit{A. sagrei} across the West Indies. While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap has enough micro-scale, local adaptive potential to contribute to reproductive isolation, should it be recruited for assortative mating.

>>>>

%\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence. Simple models of genetic drift are ruled out and we don’t think it is plasticity. However, our primary test for adaptation—looking for parallel responses across islands—also was falsified. We suggest, however, that selection could be involved, and suggest several ways this might have happened.}

@@

\paragraph{No conclusive evidence for adaptation of dewlap coloration to habitat type} One of the most informative tests for adaptation is the convergence of differentiation patterns across replicate islands or localities \citep{Losos2009, Losos2011}. Previous studies have described convergent patterns of dewlap color evolution in similar environments across islands and species \citep{Thorpe2002a, Thorpe2002b}. However, the inconsistent and idiosyncratic patterns we observed reject the test of convergence across islands, suggesting that dewlap color variation between habitats cannot be predicted by habitat identity alone in \textit{A. sagrei}. Other studies have proposed that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e. UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{A. cristatellus} and \textit{A. cooki} from Puerto Rico \citep{Leal2002, Leal2004}. Although UV reflectance was commonly involved in between-habitat divergence in \textit{A. cooki} and \textit{A. cristatellus}, we found no such patterns in \textit{A. sagrei}, where instead, we found the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}. Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}. Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A. sagrei}. Habitats on different islands may also differ in aspects other than light conditions, such as densities of predators or congeners, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}. In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A. sagrei} on islands with more coexisting species of anoles. Therefore, if local adaptation occurs, it is more likely to involve components of the environment that do not encompass our broad habitat categories.

\paragraph{Sexual selection could be at play} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Andersson1994}. This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g. plasticity or random drift). Substantial levels of promiscuity in \textit{A. sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario. However, \citet{Baeckens2018} found no link between \textit{A. sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos. Another form of sexual selection is the "good genes" model, where the cue under sexual selection is an indicator of individual quality (e.g. better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}. Several studies suggest that this possibility could be the case for anoles. For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A. brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition. \citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A. sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health. Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}. However, one would expect good genes sexual selection to favor the same coloration across the archipelago, and therefore this mechanism is unlikely to have produced idiosyncratic patterns of within-island divergence.

@@

\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors such as parasite load, as mentioned above \citep{Cook2013}. The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}. Animals lack the ability to synthesize carotenoids, and those must therefore be found in the diet, while pterins are synthesized from nucleotides \citep{Goodwin1984, Hill2002, Hill2006}. However, experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A. sagrei} \citep{Steffen2010} nor in \textit{A. distichus} \citep{Ng2013}, which also has an orange-based dewlap. This makes a plastic response to differences in diet across habitats unlikely. Furthermore, developmental plasticity during the ontogeny is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}. The differences we observed could therefore be genetically based. This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A. sagrei}. % playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore

@@

\paragraph{Implications in the context of speciation} Local adaptation can be a precursor to ecological speciation, a process that may have given rise to the adaptive radiation of \textit{Anolis} lizards \citep{Harmon2003, Gavrilets2009}. Ecologically-mediated divergence of a sexual signal may be a potent path to the evolution of reproductive isolation through divergent sexual selection \citep{Reynolds2007, Servedio2011}. Evidence suggests that dewlap coloration could take this role in anoles \citep{Ng2011, Lambert2013, Geneva2015, Ng2017}, or at least that it is frequently involved in species recognition \citep{Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013, Ingram2016, Baeckens2018}. Although a correlation between dewlap coloration and reproductive isolation is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular has been capable of rapid macroevolution throughout its phylogenetic history \citep{Nicholson2007}. For example, \textit{A. conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}. We present evidence of multiple cases of within-island dewlap divergence over small geographical scales in \textit{A. sagrei} across the West Indies. While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap has enough micro-scale, local adaptive potential to contribute to reproductive isolation, should it be recruited for assortative mating.
